---
title: "Lab 7"
author: ""
date: ""
output: 
  html_document:
    css: lab7.css
    toc: true
    number_sections: false
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA,
                      fig.width=8, 
                      fig.height=6)

library(tidyverse)
library(stringr)
library(jsonlite)
library(sf)
```

## Introduction

On May 20, 2013 an extremely powerful and devastating tornado struck
Moore, Oklahoma, a city just south of Oklahoma City. The tornado was over one
mile wide and had a maximum wind speed of 210 mph. This area is a known hotbed 
for tornadic activity. In fact, the 2013 Moore tornado followed a very similar 
path as two other large tornadoes that hit the area in 1999 and 2003, 
respectively.

How would you quantify the risk of living in certain areas of the Oklahoma
City metro? This lab will get you started in answering that question with a
focus on obtaining some of the necessary data using web scraping techniques
and APIs.

<br/>

## Data

### OK tornadoes

<div class="bkgemph">
Scrape data from http://www.tornadohistoryproject.com to get all
tornadoes recorded in Oklahoma from 1998 - 2017. Do this in 
`R/get_tornadoes.R`. Save your final data object with `saveRDS()` 
in folder `data`. Below, load your data with `readRDS()`
</div>

```{r load_tornadoes}
ok_tornadoes <- readRDS("data/ok_tornadoes.rds")
```

<br/>

### OKC school district boundaries

<div class="bkgemph">
Navigate to https://data.okc.gov/portal/page/catalog. The city of Oklahoma City 
Open Data Portal has data available that gives access to service, zoning,
public safety, and school data. Obtain the school district
boundaries.

Similar to the last section, do this in `R/get_school_districts.R`.
Save your object with `saveRDS()` in folder `data`. Below, load your 
data with `readRDS()`.
</div>

```{r load_school_districts}
school_districts <- readRDS("data/school_districts.rds")

# clean list to tibble 
school_districts <- school_districts$Features
saveRDS(school_districts, "data/schools.shp")
schools <- readRDS("data/schools.shp")
```

<br/>

### OKC fire stations

<div class="bkgemph">
Navigate to https://data.okc.gov/portal/page/catalog. The city of Oklahoma City 
Open Data Portal has data available that gives access to service, zoning,
public safety, and school data. Obtain the fire stations.

Similar to the last section, do this in `R/get_fire_stations.R`.
Save your object with `saveRDS()` in folder `data`. Below, load your 
data with `readRDS()`.
</div>

```{r load_fire_stations}
fire_stations <- readRDS("data/fire_stations")

# clean data 

fire_stations <- fire_stations$Features
```

<br/>

## Analysis

*Not required, but something to think about if you were to proceed further.*

<div class="bkgemph">
How would you develop a risk metric for each school district given the 
historical tornado data and fire station data? What other data may be valuable
to you for your analysis?
</div>

First we should see if there is a relationship between number of fire stations 
in a school district to number of tornadoes. We should use the most recent years
as districting might have changed. If we wanted a more rigorous metric, we 
should obtain fire station and district information from 1998 to 2017. We can 
then understand the relationship between number of tornadoes (maybe damage) and 
some notion of number or distance of fire stations, and then use that functional
form of the regression as the risk metric. 

First, understand how many fire stations are in each district

CANT FIGURE OUT HOW TO CONVERT TO SPATIAL DATA 
```{r}

schools %>%
  select(GeometryType, Geometry, CenterLat, CenterLong) %>%
  plot(.)

class(schools)

plot(schools)

plot(nc)

schools <- st_read(school_districts, "data/school_districts.shp", driver = "ESRI Shapefile")

nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
```

